---
title: "Package nplots"
author: "Rebecca Asiimwe"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Package nplots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


__Note__: This package is meant to create ggplots of your choice - Nice!!

Saves you the time of writing all that code by allowing you specify the kind of plot you want by issuing a few argument of choice. Let's fire away!

### Loading dependencies /required libraries

```{r}
library(nplots)
library(tidyverse)
library(gapminder)
library(plotly)
```

### Running scatter plots
```{r}
nplots_scatter(gapminder, gapminder$lifeExp, gapminder$gdpPercap, xlab="lifeExp", ylab="gdpPercap", title="lifeExp Vs GdpPercap", plotly=FALSE)
```


### Running violin plots
```{r}
nplots_violin(gapminder, gapminder$gdpPercap, gapminder$continent, jitter=TRUE, order=TRUE, xlab="", ylab="gdpPercap", title="gdpPercap across contients", plotly=TRUE, alpha=0.05, hjust=0.6)
```

#----------------------------------------------
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```




Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
